---
title: "xindi_admission_dataset"
output:
  pdf_document: default
  html_document: default
date: "2022-11-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("readxl")
library("dplyr")
library("tidyr")
library("readr")
library("ggridges")
library("viridis")
library("forcats")
library("ggpubr")
library('cowplot')
library('gridExtra')
library('ggplot2')
library('ggpubr')
library('choroplethr')
library("openintro")
library('choroplethrMaps')
```

# Import dataset

## Import institution type
```{r}
Institution_type <- data.frame(read_csv("./data_folder/CollegeScorecard_Raw_Data/Most-Recent-Cohorts-Field-of-Study.csv", show_col_types = FALSE))
Institution_type <- unique(Institution_type[c("UNITID","CONTROL")]) %>% rename(
  "Institution_type" = "CONTROL"
)
```


## Import SAT score
```{r}
Import_SAT <- function(filename, year) {
  df <- data.frame(read_csv(paste("./data_folder/CollegeScorecard_Raw_Data/", filename, sep=""), show_col_types = FALSE))
  df <- df[c("UNITID","SATVRMID", "SATMTMID")] %>% rename(
  "SAT_VR" = "SATVRMID",
  "SAT_MT" = "SATMTMID") %>% mutate(
    Year = year
  ) 
  df[df == "NULL"] <- NA
  df <- df %>% filter( rowSums(is.na(df)) == 0)
  class(df$SAT_VR) = "numeric"
  class(df$SAT_MT) = "numeric"
  return <- df
}

SAT2017 <- Import_SAT("MERGED2017_18_PP.csv", 2017)
SAT2018 <- Import_SAT("MERGED2018_19_PP.csv", 2018)
SAT2019 <- Import_SAT("MERGED2019_20_PP.csv", 2019)
SAT2020 <- Import_SAT("MERGED2020_21_PP.csv", 2020)
SAT2021 <- Import_SAT("Most-Recent-Cohorts-Institution.csv", 2021)

SAT_score <- rbind(SAT2017, SAT2018, SAT2019, SAT2020, SAT2021)

class(SAT_score$UNITID) = "character"
SAT_score <- inner_join(Institution_type, SAT_score, by = c("UNITID"))
```


## Import admission information
```{r}
Admission <- data.frame(read_csv("./data_folder/Admission_Data/Admission_2017-2021.csv", show_col_types = FALSE))

Admission <- subset(Admission, select = -c(...33) )
Admission <- Admission %>% filter( rowSums(is.na(Admission)) == 0)

names(Admission) <- c("UNITID", "Institution_Name", "Applcn_T_2021", "Applcn_M_2021", "Applcn_W_2021", "Admssn_T_2021", "Admssn_M_2021", "Admssn_W_2021", "Applcn_T_2020", "Applcn_M_2020", "Applcn_W_2020", "Admssn_T_2020", "Admssn_M_2020", "Admssn_W_2020", "Applcn_T_2019", "Applcn_M_2019", "Applcn_W_2019", "Admssn_T_2019", "Admssn_M_2019", "Admssn_W_2019", "Applcn_T_2018", "Applcn_M_2018", "Applcn_W_2018", "Admssn_T_2018", "Admssn_M_2018", "Admssn_W_2018", "Applcn_T_2017", "Applcn_M_2017", "Applcn_W_2017", "Admssn_T_2017", "Admssn_M_2017", "Admssn_W_2017")

Admission <- Admission %>% 
  pivot_longer(
    cols = colnames(Admission)[-(1:2)], 
    names_to = "Type", 
    values_to = "Student_Number",
    values_transform = list(num_nests = as.integer)
  )

Admission <- Admission %>%
  separate(col = Type, into = c("Type", "Gender", "Year"), sep = "\\_")

Admission <- Admission %>% 
  pivot_wider(
    names_from = Type, 
    values_from = Student_Number
  )

class(Admission$UNITID) = "character"
Admission <- inner_join(Institution_type, Admission, by = c("UNITID"))

```
# Result relating to admission

## Number of admission according to gender and institution type in 2017-2021
```{r}

Admission_Gender <- Admission %>% filter(Gender != "T")

stack_bar_gender_population <- ggplot(Admission_Gender, aes(fill=Gender, y= Admssn, x=Year)) + 
  geom_bar(position="stack", stat="identity") +
  xlab("Year") +
  ylab("Number of Enrolled Students") +
  facet_grid(~Institution_type) +
  ggtitle("Number of Admission Men and Women 2017-2021") 
stack_bar_gender_population
```

## Admission rate according to institution type in 2017-2021
```{r}
ggplot(Admission %>% filter(Gender == "T"), aes(x=Institution_type,y=Admssn/Applcn)) +
  geom_boxplot(aes(colour = Year)) +
  ggtitle("Admission rate in different years according to Institution type") +
  labs(y="Admission rate", x="Year") +
  scale_colour_brewer(palette = "Dark2")
```

## Admission rate according to gender and institution type in 2017-2021
```{r}
options(dplyr.summarise.inform = FALSE)
Admission_rate_year_type <- Admission %>%
  filter(Gender != "T",) %>%
  group_by(Year, Institution_type, Gender) %>%
  summarise(Admission_rate = sum(Admssn)/sum(Applcn))

ggplot(Admission_rate_year_type, aes(x = Year, y = Admission_rate, fill = Gender)) +
  geom_col(position = "dodge")+
  facet_wrap(~Institution_type) +
  ggtitle("Admission rate in different years according to gender and Institution type") +
  labs(y="Admission rate", x="Year")
```

## SAT mean scores according to subjects and institution type in 2017-2021
```{r}
SAT_score_math <- SAT_score %>%
  group_by(Year, Institution_type) %>%
  summarise(SAT_math = mean(SAT_MT))

SAT_score_verble <- SAT_score %>%
  group_by(Year, Institution_type) %>%
  summarise(SAT_verble = mean(SAT_VR))

SAT_score_avg <- inner_join(SAT_score_math, SAT_score_verble, by=c("Year", "Institution_type")) %>% 
  pivot_longer(
    cols = c(SAT_math, SAT_verble), 
    names_to = "Subject", 
    values_to = "Mean_score"
  )

ggplot(SAT_score_avg, aes(x = Year, y=Mean_score)) +
  geom_line(aes(color = Institution_type)) +
  facet_wrap(~Subject) +
  ggtitle("SAT mean scores according to subjects and institution type in 2017-2021") +
  labs(y="Mean score", x="Year")
```


## other plots
```{r}
ggplot(Admission %>% filter(Gender == "T"), aes(x=Year,y=Admssn)) +
  #geom_boxplot(outlier.shape = NA) +
  geom_boxplot(aes(colour = Institution_type)) +
  ggtitle("Admission numbers in different years according to Institution type") +
  #ylim(0, 7500) +
  labs(y="Admission Numbers", x="Year")
```
```{r}
ggplot(Admission %>% filter(Gender == "T"), aes(x=Year,y=Applcn)) +
  #geom_boxplot(outlier.shape = NA) +
  geom_boxplot(aes(colour = Institution_type)) +
  ggtitle("Application numbers in different years according to Institution type") +
  #ylim(0, 7500) +
  labs(y="Application Numbers", x="Year")
```



```{r}
options(dplyr.summarise.inform = FALSE)
Admission_year_type <- Admission %>%
  filter(Gender == "T",) %>%
  group_by(Year, Institution_type) %>%
  summarise(Admssn_number = sum(Admssn))

ggplot(Admission_year_type, aes(x = Year, y = Admssn_number, fill = Institution_type)) +
  geom_col(position = "dodge")+
  ggtitle("Total admission numbers in different years according to Institution type") +
  labs(y="Admission Numbers", x="Year")
```

```{r}
options(dplyr.summarise.inform = FALSE)
Application_year_type <- Admission %>%
  filter(Gender == "T",) %>%
  group_by(Year, Institution_type) %>%
  summarise(Applcn_number = sum(Applcn))

ggplot(Application_year_type, aes(x = Year, y = Applcn_number, fill = Institution_type)) +
  geom_col(position = "dodge")+
  ggtitle("Total application numbers in different years according to Institution type") +
  labs(y="Application Numbers", x="Year")
```




```{r}
SAT_score_math <- SAT_score %>%
  group_by(Year) %>%
  summarise(SAT_math = mean(SAT_MT))

SAT_score_verble <- SAT_score %>%
  group_by(Year) %>%
  summarise(SAT_verble = mean(SAT_VR), n = n())

SAT_score_avg <- inner_join(SAT_score_math, SAT_score_verble, by="Year") %>% 
  pivot_longer(
    cols = c(SAT_math, SAT_verble), 
    names_to = "Subject", 
    values_to = "Mean_score"
  )

ggplot(SAT_score_avg, aes(x = Year, y=Mean_score)) +
  geom_line(aes(color = Subject)) +
  scale_color_manual(values = c("darkred", "steelblue")) +
  ggtitle("SAT mean scores in different years according to different subjects") +
  labs(y="Mean score", x="Year")


```




